### 2020-01-30 abstract_file.tex
mwm initial pass

In many applications,
% of machine learning and artificial intelligence, 
one must work with models that have been trained by someone else.
For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
Here, we develop and evaluate quality metrics for pretrained neural network models at scale.
%%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
Most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best,'' but that HT-SR metrics do much better.
We also find that, for models that are not well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), norm-based metrics can qualitatively fail, while HT-SR metrics still go a very good job at predicting trends in model quality.
HT-SR metrics can also be used to characterize fine-scale properties of models, such as layer-wise \emph{correlation flow} and XXX SOMETHING ELSE.

>>>>

In many applications,
% of machine learning and artificial intelligence, 
one must work with models that have been trained by someone else.
For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
Here, we present and evaluate quality metrics for pretrained neural network models at scale.
%%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best.'' 
On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-bad,'' norm-based metrics can qualitatively fail. 
We also find that HT-SR metrics do much better, quantitatively better for at discriminating good-better-best and qualitatively better at discriminating good-bad.
HT-SR metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.

### 2020-01-30 all_models.tex
comments add for mike
mwm initial pass

@@

\charles{Still preliminary, only 309 data points here}
\paragraph{Results}
\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|}
\hline
    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
\hline
$R^{2}$ (mean) & 0.63 &0.55 &0.64 &0.64 \\
$R^{2}$ (std)  & 0.34 &0.36 &0.29 &0.30 \\
\hline
$MSE$ (mean)   & 4.54 &9.62 &3.14 &2.92 \\
$MSE$ (std)    & 8.69 &23.06 &5.14 &5.00 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average log norm metrics across 5 computer vision datasets, 17 Architectures, covering 168 (out of 309) different pretrained DNNs.  We only conclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.  These results can be readily reproduced using the Google Colab notebooks accompanying thie paper\cite{notebooks}}
\label{table:results}
\end{table}

>>>>

\michael{Now we go back to the CV models and look at over 100 of them, and draw'
more broad conclusions about why alpha and the new norm is better, by looking at the MSE}

@@

\charles{Still preliminary, only 309 data points here}
\paragraph{Results}
\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|}
\hline
    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
    & $\langle\log\Vert\mathbf{W}\Vert_{F}\rangle$ & $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$ & $\langle\hat{\alpha}=\alpha\log\lambda_{max}\rangle$ & $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ \\
\hline
$R^{2}$ (mean) & 0.63 &0.55 &0.64 &0.64 \\
$R^{2}$ (std)  & 0.34 &0.36 &0.29 &0.30 \\
\hline
$MSE$ (mean)   & 4.54 &9.62 &3.14 &2.92 \\
$MSE$ (std)    & 8.69 &23.06 &5.14 &5.00 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average log norm metrics across 5 computer vision datasets, 17 Architectures, covering 168 (out of 309) different pretrained DNNs.  We only conclude regressions for architecturs with 4 or more data points \charles{and which are postively correlated with the test error?}.  These results can be readily reproduced using the Google Colab notebooks accompanying thie paper~\cite{notebooks}}
\label{table:results}
\end{table}

### 2020-01-30 background.tex
cleaning up files
mwm initial pass

Here we will cite and discuss related work, including
\cite{MM17_TR},
\cite{MM18_TR},
\cite{MM19_HTSR_ICML},
\cite{weightwatcher_package}
\cite{MM19_KDD},
\cite{MM20_SDM},
\cite{MM20_unpub_work},

>>>>

To our knowledge, there is very little work on the particular question we are addressing: namely, how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and to do so without access to training data or testing data or details of the training protocol, etc.
Our work is, however, loosely related to several other lines of work, and we briefly discuss them here.

\paragraph{Statistical mechanical theory of NNs.}

XXX.
Cite our stuff:
\cite{MM17_TR},
\cite{MM18_TR},
\cite{MM19_HTSR_ICML},
\cite{weightwatcher_package}
\cite{MM19_KDD},
\cite{MM20_SDM},
\cite{MM20_unpub_work}.
Cite also Ganguli review and maybe other stuff.

XXX.
Distinguish between what we will call a
\emph{phenomenological theory}
(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
and what can be called a 
\emph{first principles theory} 
(that is applicable to tiny things but has no hope of scaling up).
\footnote{In most areas where there are complex highly-engineered systems (beyond complex AI/ML systems), one used phenomenological theory rather than first principles theory.  For example, one does not try to solve the Schr\"odinger equation if one is interested in building a bridge or an airplane.}

\paragraph{Norm-based capacity control theory.}
XXX.  MOST IRRELEVANT, BUT LIAO AND OUR SDM ARE RELATED.  
XXX.  MAYBE BEAT ON INFINITELY WIDE OR SOMETHING ELSE.

\paragraph{Practical problems poorly addressed by theory.}
There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
Here are several.
\begin{itemize}
\item
\textbf{Lack of accuracy metrics.}
\item
\textbf{Lack of data.}
\item
XXX. LACK OF SOMETHING ELSE.
\end{itemize}
Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text, when developing self driving car systems, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

### 2020-01-30 conclusion.tex
mwm initial pass

\section{Conclusion}
\label{sxn:conc}
XXX.  CONCLUSION.

>>>>

\section{Conclusion}
\label{sxn:conc}

XXX.  PUT CONCLUSION HERE AND WEAVE IN COMMENTS FROM BELOW.

Some other comments that we need to weave into a narrative eventually after later sections are written:
\begin{itemize}
\item
GPT versus GPT2.
What happens when we don't have enough data?
This is the main question, and we can use out metrics to evaluate that, but we also get very different results for GPT versus GPT2.
\item
The spectral norm is a regularizer, used to distinguish good-better-best, not a quality metric.
For example, it can ``collapse,'' and for bad models we can have small spectral norm.
So, it isn't really a quality metric.
\item
One question that isn't obvious is whether regularization metrics can be used as quality metrics.
One might think so, but the answer isn't obviously yes.
We show that the answer is No.
A regularizer is designed to select a unique solution from a non-unique good-better-best.
Quality metrics can also distinguish good versus bad.
\item
(We should at least mention this is like the statistical thing where we evaluate which model is better, as oposed to asking if a given model is good, I forget the name of that.)
\item
There are cases where the model is bad but regularization metric doesn't tell you that.
Quality should be correlated in an empirical way.
Correlated with good-better-best; but also tell good-bad.
\item
Question: why not use regularier for quality?
Answer: A regularizer selects from a given set of degenerate models one which is nice or unique.
It doesn't tell good versus bad, i.e., whether that model class is any good.
\item
Thus, it isn't obvious that norm-based metrics should do well, and they don't in general.
\item
We give examples of all of these: bad data; defective data; and distill models in a bad way.
(Of course, bad data means bad model, at least indirectly, since the quality of the data affects the properties of the model.)
\item
We can select a model and change it, i.e., we don't just do hyperparameter fiddling.
\end{itemize}

%Perhaps our most improtant contribution, however, is just doing things different, etc. ...

### 2020-01-30 cv_models.tex
mwm initial pass

\section{Comparison of 3 Computer Vision Models}
\label{sxn:cv}

\michael{We set up the methdology we will use later, and give examples on the CV models
and explain how to interpret things a bit. We study pretrained CV models.
We look at correlation with the reportest test accuarcies.
 We will run linear regressions. We will compute the RMSE.  
We study alpha and the 4 norms.  we could add softrank with some work.

We introduce the notion of Correlation Flow and 
tell a nice story about residual connections
with VGG RseNet and DenseNet}

\paragraph{Empirical Metrics vs Test Accuracies for CV models}

\begin{figure}[t]
    \centering
    \subfigure[ Frobenius Norm ]{
        \includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
        \label{fig:vgg-fnorm}
    }
    \qquad
    \subfigure[ Spectral Norm ]{
        \includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
        \label{fig:vgg-snorm}
    }
    \qquad
    \subfigure[ Weighted Alpha ]{
        \includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
        \label{fig:vgg-walpha}
    }
    \qquad
    \subfigure[ Alpha-Norm ]{
        \includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
        \label{fig:vgg-pnorm}
    }
    \caption{Comparison of norm metrics vs reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  Plots will be updated and replace }

    \label{fig:vgg-metrics}
\end{figure}

VGG works remarkably well

ResNet is also correlated, but the RMSE is much larger, and 2/5 of the models are almost outliers
But note that the PyTorch distribution only has a few models.
The ResNet results look better when considering all ResNet models, trained on ImageNet1K.

ADD PLOTS FOR

@@

    \subfigure[ ResNet ]{
        \includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
        \label{fig:resnet-accuracy}
    }
    \qquad
    \subfigure[ ResNet-1K ]{
        \includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
        \label{fig:resnet1k-accuracy}
    }
    \qquad
    \subfigure[ DenseNet ]{
        \includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
        \label{fig:densenet-accuracy}
    }
    \caption{$\alpha$-Norm vs reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models}
    \label{fig:cv2-accuracy}
\end{figure}

TODO:  add table of metrics for these 3 models, show that alpha-Norm is best again!

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|c|}
\hline
   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
 Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{\infty}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
\hline
 VGG & 6 & 0.56 & 0.53 & 0.48 & 0.42  \\
 ResNet & 5 & 0.9 & 1.4 & 0.61 & 0.66  \\
 ResNet-1K & 19 & 2.4 & 3.6 & 1.8 & 1.9  \\
 DenseNet & 4 & 0.3 & 0.26 & 0.16 & 0.21 \\
\hline
\end{tabular}
\end{center}
\caption{RMSE for Linear Fits of Metric to Reported Top1 Test Error for all pre-trained models in the architecture series.  All models trained on ImageNet except ResNet-1K, which was trained on ImageNet-1K. }
\label{table:models}
\end{table}

DISCUSS TABLE

\paragraph{Correlation Flow in CV Models}

Explain WHAT IS \emph{Correlation Flow}

Compare VGG, ResNet, and DenseNet in the context of how many connections that have

@@

    \subfigure[ VGG ]{
        \includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
                \label{fig:vgg-alpha-layers}
    }
    \qquad
    \subfigure[ ResNet ]{
        \includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
        \label{fig:resnet-alpha-layer}
    }
    \qquad
    \subfigure[ DenseNet ]{
        \includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
        \label{fig:densenet-alpha-layer}
    }
    \caption{Power law exponent $\alpha$ vs layer for VGG, ResNet, and DenseNet models}
    \label{fig:vgg-alpha-layers}
\end{figure}

>>>>

\section{Comparison across series of CV models}
\label{sxn:cv}

In this section, we examine the quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pyTorch}.
This also includes results for a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet1k}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.
We use our quality metrics to understand the gross properties of these models, comparing and contrasting the three (of four, if ResNet-1K is a distinct series) model series.
We also use our metrics to understand more fine-scale properties of these models as a function of model depth.
\michael{And mention distillation, if we include that in this section.}

\paragraph{Quality Metrics versus Reported Test Accuracies.}

We have examined the performance of the four quality (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
To start, Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus the reported Test accuracies~\cite{pyTorchVgg}, as well as a basic linear regression line, 
%with the Root Mean Squared Error (RMSE) shown, 
for the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BatchNormalization.
\michael{These are all log norms, right?}
\michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
All four metrics correlate quite well with the reported Top1 Accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
Similar results (not shown) are obtained for the Top5 Accuracies.  
Notice that the log $\alpha$-Norm metric $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ 
\michael{XXX IS THAT A TYPO} 
performs best, with an RMSE of $0.34$; and the second best metric is $\hat\alpha$, which is an approximate log $\alpha$-Norm metric.
\michael{Are those numbers stale, I don't see them, and some things seem in the other order.}
\michael{We should probably point to the table if we cite quantitative results.}

Figure~\ref{fig:cv2-accuracy} plots the best performing metric ($\alpha$-Norm) for the full ResNet, ResNet-1K, and DenseNet series.
(A much more detailed set of plots---including the behavior of all four metrods on each of the series---are availale at \michael{give web link}.)
For a more visual depiction, Figure~\ref{fig:cv2-accuracy} displays the log $\alpha$-Norm metric  $\log\Vert\mathbf{W}\Vert)_{\infty}^{\infty}$ versus test accuracies for the ResNet, ResNet-1K, and DenseNet series. 
Notice that ResNet series, which has been trained on the full
ImageNet dataset, has an RMSE of $0.66$, whereas the ResNet-1K series, which has been trained on the much smaller ImageNet-1K dataset,
also correlated wells, but has a much larger RMSE of $1.9$.

Table \ref{table:cv-models} summarizes results for all four metrics (Frobenius norm, Spectral norm, Weighted Alpha, and $\alpha$-Norm) applied to the VGG, ResNet, ResNet-1K, and DenseNet series.
\michael{XXX.  AGAIN, WHERE IS RMSE, IN THAT TABLE?}

For all series, all four metrics correlate quite well with the reported Top1 test accuracies.  
Moreover, the $\alpha$-Norm and/or the approximate Weighted Alpha metrics perform best.

\begin{figure}[t]
    \centering
    \subfigure[ Frobenius Norm ]{
        %\includegraphics[width=5cm]{img/VGG_lognorm_accs.png}
        \includegraphics[width=3.7cm]{img/VGG_lognorm_accs.png}
        \label{fig:vgg-fnorm}
    }
    %\qquad
    \subfigure[ Spectral Norm ]{
        %\includegraphics[width=4.9cm]{img/VGG_spectralnorm_accs.png}
        \includegraphics[width=3.7cm]{img/VGG_spectralnorm_accs.png}
        \label{fig:vgg-snorm}
    }
    %\qquad
    \subfigure[ Weighted Alpha ]{
        %\includegraphics[width=4.9cm]{img/VGG_alpha_weighted_accs.png}
        \includegraphics[width=3.7cm]{img/VGG_alpha_weighted_accs.png}
        \label{fig:vgg-walpha}
    }
    %\qquad
    \subfigure[ Alpha-Norm ]{
        %\includegraphics[width=4.9cm]{img/VGG_logpnorm_accs.png}
        \includegraphics[width=3.7cm]{img/VGG_logpnorm_accs.png}
        \label{fig:vgg-pnorm}
    }
    \caption{Comparison of quality metrics versus reported test accuracy for pretrained VGG models, trained on ImageNet, available in pyTorch.  
             XXX Plots will be updated and replaced. 
             \michael{Figures are unreadable, can we make fonts bigger, etc.}
            }
    \label{fig:vgg-metrics}
\end{figure}

@@

    \subfigure[ ResNet ]{
        %\includegraphics[width=4.2cm]{img/ResNet_logpnorm_accs.png}
        \includegraphics[width=3.7cm]{img/ResNet_logpnorm_accs.png}
        \label{fig:resnet-accuracy}
    }
    \qquad
    \subfigure[ ResNet-1K ]{
        %\includegraphics[width=4.5cm]{img/ResNet-1K_logpnorm_accs.png}
        \includegraphics[width=3.7cm]{img/ResNet-1K_logpnorm_accs.png}
        \label{fig:resnet1k-accuracy}
    }
    \qquad
    \subfigure[ DenseNet ]{
        %\includegraphics[width=4.4cm]{img/DenseNet_logpnorm_accs.png}
        \includegraphics[width=3.7cm]{img/DenseNet_logpnorm_accs.png}
        \label{fig:densenet-accuracy}
    }
    \caption{$\alpha$-Norm versus reported Top1 Error for  ResNet, ResNet-1K, and DenseNet models.}
    \label{fig:cv2-accuracy}
\end{figure}

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1in}|c|c|c|c|c|}
\hline
   &    & Frobenius Norm & Spectral Norm & Weighted Alpha & Alpha-Norm \\
 Series & \#Models   & $\Vert\mathbf{W}\Vert_{F}$ & $\Vert\mathbf{W}\Vert_{2}$ & $\hat{\alpha}=\alpha\log\lambda_{max}$ & $\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}$ \\
\hline
 VGG & 6 & 0.56 & 0.53 & 0.48 & 0.42  \\
 ResNet & 5 & 0.9 & 1.4 & 0.61 & 0.66  \\
 ResNet-1K & 19 & 2.4 & 3.6 & 1.8 & 1.9  \\
 DenseNet & 4 & 0.3 & 0.26 & 0.16 & 0.21 \\
\hline
\end{tabular}
\end{center}
\caption{RMSE for linear fits (smaller is better) of quality metrics to Reported Top1 Test Error (\cite{XXX-GIVE-REFS-AGAIN}\michael{Give refs again.}) for all pretrained models in the architecture series.  VGG, ResNet, and DenseNet were pretrained on ImageNet, and ResNet-1K was pretrained on ImageNet-1K. \michael{Do we want Frob norm squared.} }
\label{table:cv-models}
\end{table}

\paragraph{Correlation Flow in CV Models.}

We can learn much more about a pretrained model by going beyond average values to examining quality metrics for each layer weight matrix, $\mathbf{W}$, as a function of depth (or layer id).  % in the network. 
The most interesting results are seen when we plot the PL exponent, $\alpha$, as a function of depth.
%
See Figure~\ref{fig:vgg-alpha-layers}, which plots $\alpha$ for each layer (the first corresponds to data, the last to labels) for the least accurate (shallowest) and most accurate (deepest) model in each of the VGG (without BatchNormalization), ResNet-1K, and DenseNet~series.

@@

    \subfigure[ VGG ]{
        \includegraphics[width=4.5cm]{img/VGG_fnl_alpha_depth.png} 
                \label{fig:vgg-alpha-layers}
    }
    \qquad
    \subfigure[ ResNet-1K ]{
        \includegraphics[width=4.5cm]{img/ResNet_fnl_alpha_depth.png} 
        \label{fig:resnet-alpha-layer}
    }
    \qquad
    \subfigure[ DenseNet ]{
        \includegraphics[width=4.5cm]{img/DenseNet_fnl_alpha_depth.png} 
        \label{fig:densenet-alpha-layer}
    }
    \caption{PL exponent $\alpha$ versus layer for VGG, ResNet-1K, and DenseNet models.  Y axes on each plot are different.  ResNet-1K exhibits very different and much more stable behavior, which we interpret in the text as a \emph{correlation flow} across the layers.  }
    \label{fig:vgg-alpha-layers}
\end{figure}

In the VGG models, $\alpha$ systematically increases as we move down the network, from data to labels, in the Conv2D layers, starting with $\alpha\lesssim 2.0$ and reaching all the way to $\alpha\sim 5.0$; and then, in the last three, large, fully-connected (FC) layers, $\alpha$ stabilizes back down to $\alpha\in[2,2.5]$.
This is seen for all the VGG models (again, only the shallowest and deepest are shown in this figure), indicating that the main effect of increasing depth is to increase the range over which $\alpha$ can increase, thus leasing to larger $\alpha$ values in later Conv2D layers of the VGG models.
This is quite different than the behavior of either the ResNet-1K models or the DenseNet models.

For the ResNet-1K models, $\alpha$ also increases in the last few layers, but as the models get deeper, there is a wide range over which $\alpha$ values tend to remain small.
\michael{Can we say something about exceptions which are larger than for VGG or DenseNet.}
This is seen for other models in the ResNet-1K series, but it is most pronounced for the larger ResNet-1K (152) model, where $\alpha$ remains relatively stable at $\alpha\sim 2.0$, from the earliest layers all the way until we reach close to the final layers.

For the DenseNet models, $\alpha$ tends to increase as the layer id increases, in particular for layers toward the end.
This is similar to what is seen in the VGG models, but with the DenseNet models, the variance is much larger (in particular for the earlier and middle layers, where it can range all the way to $\alpha\lesssim 8.0$) and much less systematic.

We can understand Figure~\ref{fig:vgg-alpha-layers} in terms of the \emph{flow of correlation}, as follows.
Informally, one would expect a DNN model to perform well when it fascilitates the propagation of information/features across layers.
In the absence of training/test data, we can to try to quantify this by measuring the PL properties of weight matrices.
Smaller $\alpha$ values correspond to layers in which correlations across multiple scales are better captured~\cite{MM18_TR,SornetteBook}, and thus we expect that values of $\alpha$ that are stable across multiple layers enable better \emph{correlation flow} through the network.

Consider the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
VGG, while a good model in many ways, was limited by needing a massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.  
The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
ResNet greatly improved on VGG by introducing residual connections, allowing for greater accurary with far fewer parameters.
(Indeed, ResNet models of up to 1000 layers have been trained.) 
We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
This should also be contrasted with the DenseNet model, which contains many connections between every layer.
Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting the correlation flow across layers, and leaving many layers very poorly optimized.

\paragraph{Behavior on Less Thoroughy Studied CV Models.}

Figure~\ref{fig:vgg-metrics} and Table~\ref{table:cv-models} show that---for well-trained and widely-studied models---the weighted $\hat{\alpha}$ metric performs better than but similar to norm-based metrics.
This suggests an obvious question: is the $\hat{\alpha}$ metric just a variation of these more familiar norm-based metrics, or does it capture something different?  
To show that it is not just a variation and that it does capture something different, consider the following less thoroughy studied model, a compressed DNN model~\cite{CWZZ17_TR}, where we observed that the average (Frobenius or Spectral) norm increases with decreasing test error, whereas the average $\alpha$ decreases, as expected.

\begin{figure}[t]
   \centering
   \subfigure[$\lambda_{max}$ for ResNet20 layers]{
      %\includegraphics[scale=0.14]{img/resnet4d_maxev.png}
      %\includegraphics[width=3.7cm]{img/resnet4d_maxev.png}
      \includegraphics[width=4.5cm]{img/resnet4d_maxev.png}
      \label{fig:resnet204Dalpha}
   }
   \qquad
   \subfigure[$\alpha$ for ResNet20 layers]{
      %\includegraphics[scale=0.14]{img/resnet4d_alphas.png}
      %\includegraphics[width=3.7cm]{img/resnet4d_alphas.png}
      \includegraphics[width=4.5cm]{img/resnet4d_alphas.png}
      \label{fig:resnet204Dmaxev}
   }
   \caption{
     Analysis of ResNet20, distilled with Group Regularization, as implemented in the \texttt{distiller} (4D\_regularized\_5Lremoved) pre-trained models.  
     Comparison of individual layer $\mathbf{W}_{l}$ maximum eigenvalues ($\lambda_{max}$, or Spectral Norms) and  
     PL exponent $\alpha$, between baseline (green) and fine-tuned (red)  pre-trained models.   
           }
   \label{fig:resnet204D5L}
\end{figure}

Consider average metrics measured on ResNet20, trained on CIFAR10, before and after applying the Group Regularization technique, as implemented in the \texttt{distiller} package.%
\footnote{For details, see \url{https://nervanasystems.github.io/distiller/\#distiller-documentation} and also \url{https://github.com/NervanaSystems/distiller}.}
We analyze the available pre-trained 4D\_regularized\_5Lremoved baseline and fine-tuned models.  %\cite{distiller_repo}.
Figure~\ref{fig:resnet204D5L} presents the Spectral Norm ($\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}_{l}$.%
\footnote{We only include layer matrices or feature maps with $M\ge50$.}
The reported baseline test accuracies ($Top1=91.450$ and $Top5=99.750$) are better than the reported fine-tuned test accuracies ($Top1=91.020$ and $Top5=99.670$).
Thus, the results of Figure~\ref{fig:vgg-metrics} and Figure~\ref{fig:cv2-accuracy} might suggest that the baseline Spectral (and Frobenius) Norm should be \emph{smaller} than those of the layers in the fine-tuned model.
In both cases (Frobenius norm results not shown), we observe the opposite.
On the other hand, the $\alpha$ values do not systematically differ between the baseline and fine-tuned models.
Also (not shown), the average (unweighted) baseline $\alpha$ is smaller than the fine-tuned average (as would be predicted by HT-SR Theory, on which $\hat{\alpha}$ is based).

The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.  CLARIFY.}
